---
title: "Project 4 Group 6 Report"
author: "Kristen Akey, Citina Liang, Rui Liang, Mubai Liu, Wen Yin"
date: "December 2, 2020"
output:
  pdf_document:
    toc: yes
header-includes:
  - \usepackage{xcolor}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F,
                      options(width = 60),
                      matrix(runif(100), ncol = 20))
```


# Causal Inference Algorithms Evaluation

```{r load libraries}
# Load libraries
pack <- c("readr", "tidyverse", "glmnet", "pryr", "flextable")

# if package not already installed, install, and load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(pack)

for (package in pack) {
  pacman::p_load(package, character.only = TRUE, dependence=TRUE)
}
```

```{r read in data}
# Load data
lowDim <- read_csv('../data/lowDim_dataset.csv')
highDim <- read_csv('../data/highDim_dataset.csv') 
```


## Project Overview

In this project, we evaluate three causal inference algorithms. The three models include inverse propensity weighting (IPW) + L1 penalized logistic regression, regression estimate, and weighted regression + L1 penalized logistic regression. We compute the average treatment effect effect using these algorithms on two distinct datasets. 

This report includes a description of each of the algorithms, code to reproduce our results, and a final comparison of each of these models.


## Model 1: Inverse Propensity Weighting and L1 Penalized Logistic Regression

The propensity score is predicted through a lasso logistic regression model. Weights for the propensity score are then calculated using the inverse probability of treatment weighting (IPTW) formula. The average treatment effect is then calculated using inverse propensity weighting.
```{r prep data}
# Split into x, A and y
hY<-highDim$Y
hA<-highDim$A
hX<-highDim%>% select(-Y, -A) %>% as.matrix

lY<-lowDim$Y
lA<-lowDim$A
lX<-lowDim%>% select(-Y, -A) %>% as.matrix
```

```{r IPW}
# Setting alpha = 1 implements lasso regression
set.seed(0)
lasso_hd <- cv.glmnet(hX, hA,family = "binomial", alpha = 1)
lasso_ld <- cv.glmnet(lX, lA, family = "binomial",alpha = 1)

IPW<-function(x,A,model,data){
      start_time <- Sys.time()
      # Calculate the propensity score 
      lasso_model <- glmnet(x, A, alpha = 1, family = "binomial",lambda = model$lambda.min)
      propensity <- predict(lasso_model, x, type = "response")
      
      # Calculate the weights
      weight <- 1 / propensity * A + 1 / (1 - propensity) * (1 - A)
      
      resampled_data <- data %>%
            mutate(propensity = propensity,
                   weight = weight,
                   Y_Weight = Y*weight)
      
      ATE<-1/nrow(resampled_data)*(sum(resampled_data[resampled_data$A==1,"Y_Weight"])
                                   -sum(resampled_data[resampled_data$A==0,"Y_Weight"]))
      end_time <- Sys.time()
      return(list(ATE=ATE,running_time = end_time - start_time))
}
ATE_highDim<-IPW(hX,hA,lasso_hd,highDim) #ATE: -2.21809, runtime:0.02583098
ATE_lowDim<-IPW(lX,lA,lasso_ld,lowDim) #ATE: 2.21036, runtime:0.005035877

matrix(c(ATE_highDim$ATE,ATE_lowDim$ATE,
         ATE_highDim$running_time,ATE_lowDim$running_time),
         nrow = 2,byrow = TRUE,
         dimnames = list(c("ATE","running_time (secs)"), c("highDim","lowDim")))
```


\newpage
## Model 2: Regression Estimate

The regression estimate does not use the propensity score. The model creates predictions for the control and treatment groups, and calculates the average treatment effect using these predications.  

```{r prep data reg estimate}
df_ld <- lowDim %>% mutate(A = factor(A))
df_hd <- highDim %>% mutate(A = factor(A))
```


```{r}
RE <- function(df){
  # simple regression estimate
  # separate X and Y, will be used in predict function
  df_X <- df %>% select(-Y, -A)
  
  start <- Sys.time()
  # m0
  m0 <- glm(Y ~ ., data = subset(df[df$A==0,], select = -A))
  # m1
  m1 <- glm(Y ~ ., data = subset(df[df$A==1,], select = -A))
  # prediction using non-treatment model params
  Y_pred_0 <- predict(m0, newdata = df_X)
  # prediction using treatment model params
  Y_pred_1 <- predict(m1, newdata = df_X) 
  # add predicted y to the dataframe
  df <- df %>% mutate(Y_pred1 = Y_pred_1, Y_pred0 = Y_pred_0)
  
  # calculate ATE
  n <- nrow(df)
  ATE = 1/n * sum(df$Y_pred1 - df$Y_pred0)
  end <- Sys.time()
  runtime = end - start
  return(list(ATE = ATE,
              runtime = runtime))
}

matrix(c(RE(df_hd)$ATE,RE(df_ld)$ATE,
         RE(df_hd)$runtime,RE(df_ld)$runtime),
         nrow = 2,byrow = TRUE,
         dimnames = list(c("ATE","running_time (secs)"), c("highDim","lowDim")))
```


\newpage
## Model 3: Weighted Regression and L1 Penalized Logistic Regression 

The propensity score is predicted through a lasso logistic regression model. Weights for the propensity score are then calculated using the inverse probability of treatment weighting (IPTW) formula. The average treatment effect is then calculated using a weighted regression model.

```{r weighted reg. low dim}
set.seed(0)
start_time <- Sys.time()
X_low <- df_ld %>% select(-Y, -A) %>% as.matrix
A_low <- df_ld %>% select(A) %>% as.matrix
cv_l1 <- cv.glmnet(X_low, A_low, family = "binomial", alpha = 1)
l1_low <- glmnet(X_low, A_low, family = "binomial", 
                 alpha = 1, lambda = cv_l1$lambda.min)
propen_score_low <- predict(l1_low, X_low, type = "response")

# Finding weights
weight_low <- cbind(as.numeric(A_low), propen_score_low) %>% 
  as_tibble %>%
  mutate(weights = (V1/s0 + (1-V1)/(1-s0))) %>%
  select(weights)

# Linear regression for selecting covarites
filter_low <- summary(lm(Y~., data = df_ld))$coef[,4][3:24]<0.05
Z_low <- cbind(A_low, X_low[,filter_low])

Z_low <- Z_low %>% apply(2, as.numeric)

# Final Regression for ATE
Y_low <- df_ld$Y

weighted_low <- lm(Y_low ~ Z_low, weights = as.numeric(unlist(weight_low)))
ATE_low <- coef(weighted_low)[2]

end_time <- Sys.time()
running_time_low = end_time - start_time
```


```{r}
set.seed(0)
start_time <- Sys.time()
X_high <- df_hd %>% select(-Y, -A) %>% as.matrix
A_high <- df_hd %>% select(A) %>% as.matrix
cv_l1_high <- cv.glmnet(X_high, A_high, family = "binomial", alpha = 1)
l1_high <- glmnet(X_high, A_high, family = "binomial", 
                  alpha = 1, lambda = cv_l1_high$lambda.min)
propen_score_high <- predict(l1_high, X_high, type = "response")

weight_high <- cbind(as.numeric(A_high), propen_score_high) %>%
  as_tibble %>%
  mutate(weights = (V1/s0) + (1-V1)/(1-s0)) %>%
  select(weights)

filter_high <- summary(lm(Y~., data = df_hd))$coef[,4][3:ncol(X_high)]<0.05
Z_high <- cbind(A_high, X_high[,filter_high])

Z_high <- Z_high %>% apply(2, as.numeric)

Y_high <- df_hd$Y

weighted_high <- lm(Y_high ~ Z_high, weights = as.numeric(unlist(weight_high)))
ATE_high <- coef(weighted_high)[2]
end_time <- Sys.time()
running_time_high = end_time - start_time
```

```{r}
matrix(c(ATE_high,ATE_low,
         running_time_high,running_time_low),
         nrow = 2,byrow = TRUE,
         dimnames = list(c("ATE","running_time (secs)"), c("highDim","lowDim")))
```


## Model Comparisons


```{r, echo=F}
true_ate <- data.frame(
  `Low Dim` = 2.5,
  `High Dim` = -3
)

myft_true_ate<- flextable(true_ate)
ft_true <- add_header_lines(myft_true_ate, 
        values = c("True ATE") ) 
ft_true <- theme_box(ft_true)
ft_true <- fontsize(ft_true, size = 13, part='header')
ft_true <- fontsize(ft_true, size = 12, part='body')
ft_true <- align(ft_true, align = 'center', part='all')
ft_true <- autofit(ft_true)
ft_true
```

Performance = square difference of true ATE and estimated ATE

Run time (in seconds)

```{r, echo=F}
high_results <- data.frame(
            Model = c("IPW + Lasso Logistic Regression", "Regression Estimate",
                     "Weighted Regression + Lasso Logistic Regression"),
            `ATE` = c(ATE_highDim$ATE, RE(df_hd)$ATE, ATE_high[[1]]),
           `Run Time` = c(ATE_highDim$running_time[[1]], 
                          RE(df_hd)$runtime[[1]], running_time_high[[1]]),
           `Performance` = c(sqrt(abs(-3 - ATE_highDim$ATE)), 
                             sqrt(abs(-3 - RE(df_hd)$ATE)),
                            sqrt(abs(-3 - ATE_high[[1]])))
           )

low_results <- data.frame(
              Model = c("IPW + Lasso Logistic Regression", "Regression Estimate",
                     "Weighted Regression + Lasso Logistic Regression"),
            `ATE` = c(ATE_lowDim$ATE, 
                                     RE(df_ld)$ATE, ATE_low[[1]]),
           `Run Time` = c(ATE_lowDim$running_time[[1]], 
                          RE(df_ld)$runtime[[1]], running_time_low[[1]]),
              `Performance` = c(sqrt(abs(2.5 - ATE_lowDim$ATE)), 
                             sqrt(abs(2.5 - RE(df_ld)$ATE)),
                            sqrt(abs(2.5 - ATE_low[[1]])))
           )

```


```{r echo=F}
myft_high <- flextable(high_results)
ft_high <- add_header_lines(myft_high, 
        values = c("High Dimension Dataset") ) 
ft_high <- fontsize(ft_high, size = 14, part='header')
ft_high <- fontsize(ft_high, size = 10, part='body')
ft_high <- align(ft_high, align = 'center', part='all')
ft_high <- autofit(ft_high)
ft_high <- theme_box(ft_high)

myft_low <- flextable(low_results)
ft_low <- add_header_lines(myft_low,
        values = c("Low Dimension Dataset") )
ft_low <- fontsize(ft_low, size = 14, part='header')
ft_low <- fontsize(ft_low, size = 10, part='body')
ft_low <- align(ft_low, align = 'center', part='all')
ft_low <- autofit(ft_low)
ft_low <- theme_box(ft_low)

ft_low
ft_high
```







