---
title: "Regression Estimate"
author: Citina Liang
output: pdf_document
---

### Some notes from the literature review
* In the regression estimate in Chan et al. (2010), it calculates the ATT (Average Treatment Effect on Treatment), where the expression is $\hat{\Delta}_{reg}=n_1^{-1}\sum\limits_{exposed}(\hat{m}_1(\mathbf{X}_i)-\hat{m}_0(\mathbf{X}_i))$

* Differently, our goal is to estimate the ATE (Average Treatment Effect) and 95% confidence interval. The ATE formula is $\hat{\Delta}_{reg}=N^{-1}\sum\limits_{i=1}^{N}(\hat{m}_1(\mathbf{X}_i)-\hat{m}_0(\mathbf{X}_i))$

* Propensity scores can be used to choose matched control group that is more similar with the trt group to calculate more accurate Y

* Correlation doesn't imply causation, but we can assume that to test if a treatment make a difference in result variable. One problem is we don't have all variables that affect the outcome, so we have un-confoundedness assumption

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# load libraries
library(readr)
library(tidyverse)
```

```{r}
# read in data
df_ld <- read_csv('../data/lowDim_dataset.csv') #475  24
df_hd <- read_csv('../data/highDim_dataset.csv') #2000  187
df_ld <- df_ld %>% mutate(A = factor(A))
df_hd <- df_hd %>% mutate(A = factor(A))

# seperate X and Y, will be used in predict function
df_ld_X <- df_ld %>% select(-Y, -A)
df_hd_X <- df_hd %>% select(-Y, -A)

# data description
df_ld %>% head()
df_hd %>% head()

df_ld$A %>% table() %>% prop.table()
df_hd$A %>% table() %>% prop.table()

df_ld %>% group_by(A) %>% summarize(mean(Y))
df_hd %>% group_by(A) %>% summarize(mean(Y))
```

```{r}
# simple regression estimate
# m0
m_ld0 <- glm(Y ~ ., data = subset(df_ld[df_ld$A==0,], select = -A))
m_hd0 <- glm(Y ~ ., data = subset(df_hd[df_hd$A==0,], select = -A))
# m1
m_ld1 <- glm(Y ~ ., data = subset(df_ld[df_ld$A==1,], select = -A))
m_hd1 <- glm(Y ~ ., data = subset(df_hd[df_hd$A==1,], select = -A))

# prediction using non-treatment model params
Y_pred_ld0 <- predict(m_ld0, newdata = df_ld_X)
Y_pred_hd0 <- predict(m_hd0, newdata = df_hd_X)

# prediction using treatment model params
Y_pred_ld1 <- predict(m_ld1, newdata = df_ld_X)
Y_pred_hd1 <- predict(m_hd1, newdata = df_hd_X)

# add predicted y to the dataframe
df_ld <- df_ld %>% mutate(Y_pred1 = Y_pred_ld1, Y_pred0 = Y_pred_ld0)
df_hd <- df_hd %>% mutate(Y_pred1 = Y_pred_hd1, Y_pred0 = Y_pred_hd0)

par(mfrow=c(2,3))
plot(df_ld$A, Y_pred_ld1)
plot(df_ld$A, Y_pred_ld0)
plot(df_ld$A, df_ld$Y)
plot(df_hd$A, Y_pred_hd1)
plot(df_hd$A, Y_pred_hd0)
plot(df_hd$A, df_hd$Y)

# calculate ATE
ATE <- function(df){
  n <- nrow(df)
  return(1/n * sum(df$Y_pred1 - df$Y_pred0))
}
ATE(df_ld) #2.526944
ATE(df_hd) #-2.95978
```